taxadb: A High-Performance Local Taxonomic Database Interface

### 2018-12-28 paper/manuscript.Rmd
various exploration:
looking a BBS data example for manuscript.
need to deal with fuzzy matching

>>>>

```{r}
## Consider: Working from a focal authority:
## select all rows where focal authority did not match but another authority did
## for each matched authority, access species name associated with id, and requery

## OR consider: access accepted species name associated with any authority, requery that in original authority?
##

## OR consider? generate a combined taxonid table for all authorities
```

```{r}
bbs_ids %>% 
  select(-species) %>% 
  purrr::map_dbl(function(x) sum(!is.na(x)))
```

ITIS resolves the most names, but no authority resolves all the names provided.

```{r}
unmatched <- species_list[is.na(bbs_ids$itis)]

first_two <- function(x){ tmp <- strsplit(x, " ")[[1]]; paste(tmp[[1]], tmp[[2]]) }
drop_sp <- function(x) gsub("(\\w+)\\s+sp\\.", "\\1", x)

matched <- unmatched %>% 
  map_chr(first_two) %>% 
  drop_sp() %>%
  get_ids("itis", "prefix")

matched %>% is.na() %>% sum()

still_umatched <- unmatched[is.na(matched)]

```{r}
first_two_fuzzy <- function(x){ 
  tmp <- strsplit(x, " ")[[1]]
  paste0(tmp[[1]], " ", tmp[[2]], "%") 
}

only_genus <- function(x){ 
  purrr::map_chr(x, function(x) strsplit(x, " ")[[1]][[1]])
}
name_to_regex <- function(x){ 
  x <- gsub("(\\w+)\\s+sp\\..*", "\\1.*", x)
  # x <- gsub("(\\w+)\\s+(\\w+).*", "\\1 \\2.*", x)
  x <- gsub("(\\w+)\\s+(\\w{1,3}).*", "\\1 \\2.*", x)
  x
}
name_to_regex(still_umatched)

## subset by genus first
tmp <- 
  right_join(taxa_tbl("itis", "hierarchy"), 
           data_frame(genus = only_genus(still_umatched)),
           copy = TRUE) %>% select(id) %>% 
  left_join(taxa_tbl("itis", "taxonid"), by="id") %>%
  distinct() %>% 
  collect()

df <- data_frame(regex_name = name_to_regex(still_umatched))
i <- vapply(df$regex_name, 
            function(x) which(grepl(x, tmp$name)),
            integer(1))
tmp[i,] %>%  select(id, name) %>% distinct()

tmp[which(grepl("Rallus.*", tmp$name)), ] %>%  select(id, name) %>% distinct()

```{r}
matched <- tmp %>%   
  fuzzyjoin::regex_inner_join(
    data.frame(regex_name = name_to_regex(still_umatched)),
    by = c(name = "regex_name")
  )
```

```{r}
matches <- 
  right_join(taxa_tbl("itis", "hierarchy"), 
           data_frame(genus = only_genus(species_list)),
           copy = TRUE) %>% 
  distinct() %>% 
  collect() 
# uses smaller table filtered by genus
library(fuzzyjoin)
left_join(data_frame(species = species_list), matches)

stringdist_left_join(data_frame(species = species_list), matches)

  right_join(data_frame(species = first_two(unmatched))) %>% 
  select(id, species) %>% 
  distinct()

taxa_tbl("itis", "taxonid") %>%
  filter(name %in% c("Rallus%", "Vermivora%"), rank == "species")

taxa_tbl("itis", "hierarchy") %>% filter(genus == "Rallus") %>% select(genus, species)
```

```{r}
itis <- ids(species_list, "itis") 
col <- accepted_name(bbs_ids$itis, "itis") %>% pull(name) %>% get_ids("col")
sum(!is.na(col))
```

```{r}
unmatched <- bbs_ids %>%  
  select(-species) %>%
  purrr::map_dfc(function(x) !is.na(x)) %>% 
  rowSums() < 1

sum(!unmatched)
```
